---
title: 'Estudo de Caso 2 - Avaliação e comparação do retorno médio de ações'
author: "Ana Júlia Martins, Antônio Carlos da Anunciação, Melchior Augusto Syrio de Melo"
date: "25 de novembro de 2024"
output:
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
---

```{r setup,results='hide',warning=FALSE,echo=FALSE}
# A few initial definitions just to make sure all required packages are installed. Change as needed.
# NOTE: It may echo some weird messages to the PDF on the first compile (package installation messages). Run twice and the problem will (hopefully) go away.
if (!require(ggplot2, quietly = TRUE)){
      install.packages("ggplot2")
      }
if (!require(devtools, quietly = TRUE)){
      install.packages("devtools")
      }
 if (!require(broom, quietly = TRUE)){
       devtools::install_github("dgrtwo/broom")
      }
# if (!require(GGally, quietly = TRUE)){
#       install.packages("GGally")
#       }
if (!require(dplyr, quietly = TRUE)){
      install.packages("dplyr")
      }
```
## Summary
This document provides a template for the Case Study reports. Reports should always start a short executive summary (call it an _Abstract_ if you want) to give the reader a general idea of the topic under investigation, the kind of analysis performed, the results obtained, and the general recommendations of the authors.

There are (at least) two ways to render this document to **pdf** in R: one easy and one... well, less easy.

### 1. The easy way
Use [RStudio](http://rstudio.com/) as your editor, open the **.Rmd** file and click the _Knit PDF_ button at the top of the editor.

### 2. The slightly-less-easy way
If you're using any other R editor (such as the basic [R](http://cran.r-project.org) editor), you have to use the _render()_ function from the **rmarkdown** package:

```{r, eval=F}
install.packages("devtools")                    # you only have to install it once
library(devtools)
install_github("rstudio/rmarkdown")             # you only have to install it once
library(rmarkdown)
render("report_template.Rmd","pdf_document")    # this renders the pdf
```

## Experimental design
A section detailing the experimental setup. This is the place where you will define your test hypotheses, e.g.:
$$\begin{cases} H_0: \mu = 10&\\H_1: \mu<10\end{cases}$$

including the reasons behind your choices of the value for $H_0$ and the directionality (or not) of $H_1$. 

This is also the place where you should discuss (whenever necessary) your definitions of minimally relevant effects ($\delta^*$), sample size calculations, choice of power and significance levels, and any other relevant information about specificities in your data collection procedures.

### Description of the data collection

Os dados utilizados neste estudo são referentes aos preços de fechamento de cinco ações, extraídos de um arquivo CSV. Cada arquivo contém 36 linhas e 5 colunas, onde:

- Cada linha representa o preço de fechamento mensal das ações, com a linha 1 correspondendo ao mês mais recente e a linha 36 ao mês mais distante.
- Cada coluna representa uma das cinco ações analisadas.
- O conteúdo da posição (i, j) refere-se ao preço de fechamento da ação j no mês i.

Com base nesse conjunto de dados, foi calculado o retorno mensal de cada ação utilizando a fórmula:

\[
\text{Retorno} = \frac{\text{Preço no mês atual} - \text{Preço no mês anterior}}{\text{Preço no mês anterior}}
\]

O primeiro mês de cada coluna foi descartado, pois não há um valor anterior para o cálculo do retorno. O processo resultou em uma tabela com os retornos mensais das cinco ações.

```{r loaddata}
# Ler o arquivo CSV com os preços de fechamento das ações
precos <- read.csv("DadosAcoesGrupoH.csv", header=FALSE)

# Calcular o retorno mensal de cada ação
# Retorno mensal = (Preço atual - Preço anterior) / Preço anterior
retornos <- precos
for (j in 1:ncol(precos)) {
  for (i in 2:(nrow(precos))) {
    retornos[i-1, j] <- ((precos[i, j] - precos[i-1, j])) / precos[i-1, j]
  }
}

# Remover a ultima linha, porque a primeira linha da tabela de preços foi descartada
retornos <- retornos[1:nrow(precos)-1, ]
```


## Exploratory Data Analysis

To get an initial feel for the relationships between the relevant variables of your experiment it is frequently interesting to perform some preliminary (exploratory) analysis. This is frequently referred to as _getting a feel_ of your data, and can suggest procedures (such as outlier investigation or data transformations) to experienced experimenters.

# ```{r pairs,fig.width=8,echo=TRUE,message=FALSE,fig.cap="Exploring the effect of car transmission on mpg values"}
# library(GGally,quietly = T, warn.conflicts = F) # This is just me getting fancy.
#                                                 # There are much simpler ways ;-)
# ggpairs(data=mtcars,columns=c(1,9),title="MPG by transmission type",
#         upper=list(combo="box"),lower=list(combo="facethist"),
#         diag=list(continuous="densityDiag",discrete="barDiag"))
# ```

Your preliminary analysis should be described together with the plots. In this example, two facts are immediately clear from the plots: first, **mpg** tends to correlate well with many of the other variables, most intensely with **drat** (positively) and **wt** (negatively). It is also clear that many of the variables are highly correlated (e.g., **wt** and **disp**). Second, it seems like manual transmission models present larger values of **mpg** than the automatic ones. In the next section a linear model will be fit to the data in order to investigate the significance and magnitude of this possible effect.

## Statistical Analysis
Your statistical analysis should come here. This is the place where you should fit your statistical model, get the results of your significance test, your effect size estimates and confidence intervals.

```{r fitmodel,results='hold'}
model<-aov(mpg~am*disp,data=mtcars)
summary(model)
```

### Checking Model Assumptions
The assumptions of your test should also be validated, and possible effects of violations should also be explored.
```{r resplots,fig.width=8,echo=TRUE,message=FALSE, fig.cap = "Residual plots for the anova model"}
par(mfrow=c(2,2), mai=.3*c(1,1,1,1))
plot(model,pch=16,lty=1,lwd=2)
```

### Conclusions and Recommendations
The discussion of your results, and the scientific/technical meaning of the effects detected, should be placed here. Always be sure to tie your results back to the original question of interest!